---
title: "p8105_hw1_gm2794"
author: "GMA"
date: "21 de septiembre de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Problem 1

This problem focuses on variable types and coercion.

Create a data frame comprised of:

A random sample of size 10 from a uniform[0, 5] distribution
A logical vector indicating whether elements of the sample are greater than 2
A (length-10) character vector
A (length-10) factor vector
```{r}
library(tidyverse)
dataframe1 = tibble( 
  sample = runif(10, min=0, max=5),
  logic = sample > 2,
  charac = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  fact = factor(c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"))
  )
```
Try to take the mean of each variable in your dataframe. What works and what doesn’t? Why? 
```{r}
mean (dataframe1$sample)
mean (dataframe1$logic)
mean (dataframe1$charac)
mean (dataframe1$fact)
#The two last variables do not yield mean as they are not numeric
```
In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens? 

```{r}
as.numeric(dataframe1$logic)
as.numeric(dataframe1$charac)
as.numeric(dataframe1$fact)
#The character variable does not convert into numeric
```

In a second code chunk, convert your character variable from character to factor to numeric; similarly, convert your factor variable from factor to character to numeric. What happens?

```{r}
as.numeric(as.factor(dataframe1$charac))
as.numeric(as.character(dataframe1$fact))

```

Problem 2

This problem focuses on plotting and the use of inline R code.

Create a data frame comprised of:
x: a random sample of size 1000 from a standard Normal distribution
y: a random sample of size 1000 from a standard Normal distribution
A logical vector indicating whether the x + y > 0
A numeric vector created by coercing the above logical vector
A factor vector created by coercing the above logical vector

```{r}
library(tidyverse)
dataframe2 = tibble(
  x = rnorm (1000),
  y = rnorm (1000),
  xy0 = x + y > 0,
  xy0numericvector = as.numeric(xy0),
  xy0factorvector = as.factor(xy0numericvector)
)
```

Write a short description of your vector using inline R code, including: * the size of the dataset 
* the mean and median of x 
* the proportion of cases for which the logical vector is TRUE

```{r}
dim(dataframe2)
# includes mean and median
summary(dataframe2)
```
Make a scatterplot of y vs x; color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). 
```{r}
ggplot(dataframe2, aes(x = x, y = y, color = xy0)) +
  geom_point()
```

Make a second and third scatterplot that color points using the numeric and factor variables, respectively, and comment on the color scales.

```{r}
ggplot(dataframe2, aes(x = x, y = y, color = xy0numericvector)) +
  geom_point()
# this graph uses a range of colours
```
```{r}
ggplot(dataframe2, aes(x = x, y = y, color = xy0factorvector)) +
  geom_point()
# this graph only has two options as in my first one
```
Export your first scatterplot to your project directory using ggsave.
```{r}
ggplot(dataframe2, aes(x = x, y = y, color = xy0)) +
  geom_point()

ggsave("dataframe2_plot.pdf")
```

